Real-space computation of $E$/$B$-mode maps I: Formalism and Compact Kernels

### 2017-11-02 articles/formalism/article.tex
Heavy edit of abstract

\begin{document}
\title{Quantifying the locality of operators that translate Stokes $Q$ \& $U$ to scalars $E$ \& $B$}
\title{Real space operators for Stokes $Q$ \& $U$ to scalars $E$ \& $B$ translation}
\author[]{Aditya Rotti and}
\author[]{Kevin Huffenberger}
\affiliation[]{Department of Physics, Florida State University, Keen Physics Building, 77 Chieftan Way, Tallahassee, Florida, U.S.A.}
\emailAdd{adityarotti@gmail.com}
\emailAdd{khuffenberger@fsu.edu}

\abstract{
\noindent  We derive fully sky real space operators to translate Stokes Q \& U parameters to scalars $E$ \& $B$ and vice versa. We explicitly show that these local real space operator are fully characterized by the spin-0 $Y_{\ell 2}$ spherical harmonic functions. These real space operators make transparent the association of radial and tangential patterns of polarization with the E modes and that of  the clockwise and anti-clockwise pinwheel patterns of polarization with B-modes. We cast the standard CMB polarization analysis operators in a matrix-vector notation which elucidate these derivations. Using this new notation also allows us to derive real space operators which decompose the measured Stokes parameters into those corresponding to E-modes and B-modes respectively, without ever evaluating the scalar fields themselves. We use these analytical derivations us to quantify the non-local nature of the relation between the two representations of the polarization field. We present a prescription for generalizing these operators, which allow the non-local behavior of these operators to be a tunable parameter. We derive constraints on these generalized operators to ensure reliable recovery of the standard $E$ and $B$ fields and the corresponding power spectra. \revisit{This is part 1 of a series of papers where we have derived the real space operators. In the following papers we will explore some of their applications.} \comment{Needs better blending.}}

>>>>

\begin{document}
%\title{Quantifying the locality of operators that translate Stokes $Q$ \& $U$ to scalars $E$ \& $B$}
%\title{Real space operators for Stokes $Q$ \& $U$ to scalars $E$ \& $B$ translation}
\title{Real-space computation of $E$/$B$-mode maps I: Formalism and Compact Kernels}
\author[]{Aditya Rotti and}
\author[]{Kevin Huffenberger}
\affiliation[]{Department of Physics, Florida State University, Keen Physics Building, 77 Chieftan Way, Tallahassee, Florida, U.S.A.}
\emailAdd{adityarotti@gmail.com}
\emailAdd{khuffenberger@fsu.edu}

\abstract{
  \noindent 
  \revisit{  We derive full-sky, real-space operators that convert between polarization Stokes $Q$/$U$ parameters to the coordinate-independent, scalar $E$/$B$ modes that are widely used in Cosmic Microwave Background and cosmic shear analysis.  The covolution kernels split naturally into angular and radial parts, and we show explicitly how the spatial extent of the convolution kernel depends on the targeted band-limit.  We show that a arbitrary radial dependence can produce $E$/$B$-like maps and that these are simply filtered versions of the standard $E$/$B$ maps.  This allows us to compute $E$/$B$ maps in real space with a compactly-supported kernel, an approach that can guarantee the avoidance of known foreground regions and can be employed in a massively parallel scheme at high-resolution.  We can compute power spectra using standard techniques, and recover the power spectrum of the sky with a simple window function.  
    %
     We cast the standard CMB polarization analysis operators in a matrix-vector notation which facilitates the derivations and shows that the kernels relate directly to spin-0 $Y_{\ell 2}$ spherical harmonic functions.  This new notation also allows us to derive real space operators which decompose the measured Stokes parameters into their even and odd-parity parts, without ever evaluating the scalar $E$/$B$ fields themselves.  This paper is the first part in a series of papers that explore real-space computation of polarization modes and their applications.
  } 
  \\
  \\
%  We derive fully sky real space operators to translate Stokes Q \& U parameters to scalars $E$ \& $B$ and vice versa. We explicitly show that these local real space operator are fully characterized by the spin-0 $Y_{\ell 2}$ spherical harmonic functions. These real space operators make transparent the association of radial and tangential patterns of polarization with the E modes and that of  the clockwise and anti-clockwise pinwheel patterns of polarization with B-modes. We cast the standard CMB polarization analysis operators in a matrix-vector notation which elucidate these derivations. Using this new notation also allows us to derive real space operators which decompose the measured Stokes parameters into those corresponding to E-modes and B-modes respectively, without ever evaluating the scalar fields themselves. We use these analytical derivations us to quantify the non-local nature of the relation between the two representations of the polarization field. We present a prescription for generalizing these operators, which allow the non-local behavior of these operators to be a tunable parameter. We derive constraints on these generalized operators to ensure reliable recovery of the standard $E$ and $B$ fields and the corresponding power spectra. \revisit{This is part 1 of a series of papers where we have derived the real space operators. In the following papers we will explore some of their applications.} \comment{Needs better blending.}
}
